---
title: "final"
author: "Yi-chin Tzou"
date: "2023-05-10"
output: html_document
---

```{r, warning=FALSE}
#import required packages
library(xts)
library(lubridate)
library(tseries)
library(forecast)
library(tidyverse)
library(TSA)
```

#Load data
```{r,message=FALSE}
data <-
  list.files(path = "/Users/jasmine19970120/Documents/UChicago/2023_Spring/Time_series/final", pattern = "*.csv") %>% 
  map_df(~read_csv(.))
head(data)
```

Since we want to predict the demand for taxi in New York, will only keep column taxi_id and trip_start_timestamp for further analysis
```{r}
#select taxi_id, trip_start_timestamp
df<-data[,c("taxi_id","trip_start_timestamp")]

# Convert "trip_start_timestamp" column to POSIXct
df$trip_start_timestamp <- as.POSIXct(df$trip_start_timestamp)

head(df)
```

## Group data by date and hour
```{r}
library(dplyr)
# Extract date and hour from "trip_start_timestamp" and create new columns
df$date <- as.Date(df$trip_start_timestamp)
df$hour <- format(df$trip_start_timestamp, format = "%H")

# Group by date and hour and calculate total trip count
df <- df %>%
  group_by(date, hour) %>%
  summarise(total_trips = n())

# Sort the resulting dataframe by date and hour
df <- df[order(df$date, df$hour), ]
head(df)
```

```{r}
#select data only till 2016-04-01 ~ 06-07
df <- subset(df, date >= as.Date("2016-04-01") & date <= as.Date("2016-06-07"))
head(df)
```

```{r}
#check if all dates has hour data
table(df$date)
```
From the above table, all the dates has data for hours

# convert data to time series type
```{r}
#time_index <- seq(from = as.POSIXct("2013-04-01 00:00"), 
#                  to = as.POSIXct("2013-05-07 23:00"), by = "hour")

#ts_data <- xts(df$total_trips, order.by = time_index, frequency = 24)

# Convert the date and hour columns to a POSIXct format
df$timestamp <- as.POSIXct(paste(df$date, df$hour), format = "%Y-%m-%d %H")
df
#conver total_trip to time series data
ts_data <- ts(df$total_trips, start = start(df$timestamp), frequency = 24)
ts_data
```

# Exploratory data analysis
```{r}
#plot times series data
plot(ts_data,main="Total trips per hour 4/1-6/7")
# Plot total trips by hour
ggplot(df, aes(x = hour, y = total_trips)) +
  geom_bar(stat = "identity", fill = "seagreen") +
  labs(x = "Hour", y = "Total Trips", title = "Total Trips by Hour")

```

## Perform ADF and KPSS test and conclude that time series is trend stationary
```{r}
#perform ADF test
adf.test(ts_data)
#perform KPSS test
kpss.test(ts_data, null="Trend")
```
From the above ADF test, the p-value<0.05, meaning it rejects the null hypothesis, and from KPSS test, the p-value>0.05, which fail to reject the null hypothesis, meaning that from both tests conclude the same result: the data does not have a unit root and is stationary.

# train, test split
```{r}
# Define the start and end indices for the training set
train_start <- c(1, 1)
train_end <- c(61, 24)

# Define the start and end indices for the test set
test_start <- c(62, 1)
test_end <- c(68, 24)

# Split the time series into the training and test sets
train <- window(ts_data, start = train_start, end = train_end)
test <- window(ts_data, start = test_start, end = test_end)
```

#Plot the ACF, PACF to check for seasonality and pattern in our data
```{r}
# Plot the ACF and PACF of the dataset
acf(ts_data,lag.max = 100,main="ACF of dataset")
pacf(ts_data,lag.max = 100,main="PACF of dataset")
```


## TBATS
```{r}
# Fit an TBATS model to the train data
model1 <- tbats(train)
# Print the model summary
summary(model1)

#component of the model
comp <- tbats.components(model1)
plot(comp)

# Make forecasts for the next 168hours
predict1 <- forecast(model1, h = 48)
predict1
# Plot forecast value
plot(predict1)
```

## Check the model performance
```{r}
# Check residuals from a time series model
checkresiduals(model1)
```

```{r}
# Calculate accuracy metrics
accuracy_tbats <- accuracy(predict1, test)
accuracy_tbats
```

Since from the Ljung-Box test, we will transform the data to see if the model fits better
```{r}
# Transform the data using Box Cox transformation
lambda_value <- BoxCox.lambda(train)

# Fit a TBATS model with modified parameters
model2 <- tbats(train, seasonal.periods = 24, use.box.cox = TRUE, lambda = lambda_value, use.parallel = FALSE)
#predict on next 48 hrs
predict2 <- forecast(model2, h = 48)
predict2
# Plot forecast value
plot(predict2)
```
```{r}
checkresiduals(model2)
```
```{r}
# Calculate accuracy metrics
accuracy_tbats2 <- accuracy(predict2, test)
accuracy_tbats2
```
